\Title

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motionclouds}). We develop a SNN-like method able to learn hetero-synaptic delays to perform motion detection on a synthetic event-based dataset. Because neuromorphic devices are, by design, good candidates to integrate computations with time, we highlight the fact that this event-driven algorithm is transferable to such hardware.
%
\section{Methods}
\label{sec:methods}
%
%
% We expect our model to be able to efficiently detect temporal patterns. Here, we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion. This ground truth will be used to learn the weights and delays of the network in a supervised fashion.
%
Our SNN model will be tested on an appropriate computational task of detecting motion in an event-based sensory input. By analogy with biological requirements, we will design a paradigm simulating eye movements where an input image is moved by a parameterized saccadic trajectory. The task is then to guess the movement as accurately as possible (see fig: movement - task). This is, for example, useful in real life situations to compensate for this motion and provide a motion invariant image categorization by stabilizing the image on the retina.
%
\subsection{Task definition: fast motion detection}
%
%Taking analogy with biological conditions, a common use case for motion detection is to estimate the motion generated on a sensor by saccade-like movements. Saccades are rapid movements of the eyeball which allow to reposition the axis of view in a short period. In humans, these are particularly rapid 
%To test our model, we will quantify its ability to categorize different motions. In that order, we will generate a natural
Let us first define a trajectory inspired by the biological movements of the eyes. Indeed, these movements allow us to dynamically actuate the center of vision in the field of vision. For example, saccades are rapid movements of the eyeball that reposition the center of vision towards points of interest. In humans, these are very frequent (on average every 3 seconds [ ]). They are produced very rapidly (about 80ms) and the maximum speed is proportional to the amplitude of the displacement (up to 600 deg/sec for a large saccade of 30deg amplitude). At a more microscopic scale, the gaze moves with an incessant drift similar to a Brownian-like trajectory (Poletti 2015). To maintain the full generality of the task, we defined eye movements using a form of random walk (Engbert 211). This approach first defines a defined set of possible 2D motions in polar coordinates. Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-velocity=7 different velocities (see inset of Fig: motion-task, a). Note that the velocities are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$. Next, we define the trajectory as segments whose duration is drawn from a Poisson distribution with an average block length = 20 ms - similar to a Lévy flight. Finally, the trajectory is integrated by assuming 1/ that velocities are uniformly and independently sampled from the set of different motion-sets and 2/ that motion is uniform during a time segment. The resulting instances yield trajectories qualitatively similar to those observed for eye movements as observed for example in humans (see figure fig: motion-task - a).

\begin{figure}[h!]%[ht!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/motionclouds.pdf}
    \note{TODO: redo figure according to caption}
    \caption{
    {\bf Motion detection task.} {\bf (a)} To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like Brownian motion, and we show here on example path of the trajectory. {\bf (b)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (here the cross at the center of the image) such as to reproduce the effect of displacing the eye. {\bf (c)} Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation. Mimicking the retina, this representation codes for proportional increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events. This will constitute the input to the spiking neural network.}
    \label{fig:motionclouds}
\end{figure}

Once these eye-movement trajectories are simulated, we can use them on a visual scene. For this, we selected a database of 10 grayscale natural images which are typically used for studying tu statistics of natural images [olshausen 96]. Note that these are reprocessed in order to equalize the energy in each frequency band (that is, whitened). This process is known to occur as early as the retina-thalamic pathway [dan 1996]. These images are of size 512 × 512 and we will extract sub-images of 1281 128 which will be positionned around the center of gaze at every time step (see figure fig : motion -task - b). We will discretize time in 1ms true bins and produce movies of duration N-T = 250 ms (that is, approximately corresponding to the mean duration between two saccades). To avoid border effects, we will position one draw of the full trajectory at random in image space such that the sub-image is translated using the position given by the trajectory at each time step. The translation is computed using a roll of coordinates in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space [Perrinet, 2015]. Note that the amplitude of the displacement is relative to the time bin and we defined speed such that a speed of V-1 corresponds to a motion of one pixel per frame (that is, per
time bin).

>>>>

In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motion-task}). We develop a SNN-like method able to learn hetero-synaptic delays to perform motion detection on a synthetic event-based dataset. Because neuromorphic devices are, by design, good candidates to integrate computations with time, we highlight the fact that this event-driven algorithm is transferable to such hardware.
%
\section{Methods}
\label{sec:methods}
%
%
% We expect our model to be able to efficiently detect temporal patterns. Here, we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion. This ground truth will be used to learn the weights and delays of the network in a supervised fashion.
%
Our SNN model will be tested on an appropriate computational task of detecting motion in an event-based sensory input. By analogy with biological requirements, we will design a paradigm simulating eye movements where an input image is moved by a parameterized saccadic trajectory. The task is then to guess the movement as accurately as possible (see fig: movement - task). This is, for example, useful in real life situations to compensate for this motion and provide a motion invariant image categorization by stabilizing the image on the retina.
%
\subsection{Task definition: fast motion detection}
%
%Taking analogy with biological conditions, a common use case for motion detection is to estimate the motion generated on a sensor by saccade-like movements. Saccades are rapid movements of the eyeball which allow to reposition the axis of view in a short period. In humans, these are particularly rapid 
%To test our model, we will quantify its ability to categorize different motions. In that order, we will generate a natural
Let us first define a trajectory inspired by the biological movements of the eyes. Indeed, these movements allow us to dynamically actuate the center of vision in the field of vision. For example, saccades are rapid movements of the eyeball that reposition the center of vision towards points of interest. In humans, these are very frequent (on average every 3 seconds [ ]). They are produced very rapidly (about 80ms) and the maximum speed is proportional to the amplitude of the displacement (up to 600 deg/sec for a large saccade of 30deg amplitude). At a more microscopic scale, the gaze moves with an incessant drift similar to a Brownian-like trajectory (Poletti 2015). To maintain the full generality of the task, we defined eye movements using a form of random walk (Engbert 211). This approach first defines a defined set of possible 2D motions in polar coordinates. Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-velocity=7 different velocities (see inset of \seeFig{motion-task}-(a)). Note that the velocities are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$. Next, we define the trajectory as segments whose duration is drawn from a Poisson distribution with an average block length = 20 ms - similar to a Lévy flight. Finally, the trajectory is integrated by assuming 1/ that velocities are uniformly and independently sampled from the set of different motion-sets and 2/ that motion is uniform during a time segment. The resulting instances yield trajectories qualitatively similar to those observed for eye movements as observed for example in humans (\seeFig{motion-task}-(a)).

\begin{figure}[h!]%[ht!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/motion-task.pdf}
    \note{TODO: redo figure according to caption}
    \caption{
    {\bf Motion detection task.} {\bf (a)} To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like Brownian motion, and we show here on example path of the trajectory. {\bf (b)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (here the cross at the center of the image) such as to reproduce the effect of displacing the eye. {\bf (c)} Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation. Mimicking the retina, this representation codes for proportional increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events. This will constitute the input to the spiking neural network.}
    \label{fig:motion-task}
\end{figure}

Once these eye movement trajectories are generated, we can apply them to a visual scene. For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images [olshausen 96]. Note that these are pre-processed to equalize the energy in each frequency band (i.e., whitened). This process is known to occur as early as the retino-thalamic pathway [dan 1996]. These images are 512 × 512 in size and we will extract sub-images of size 128 x 128 which will be positioned around the center of the gaze at each time step (see \seeFig{motion-task}-(b)). We will discretize the time in 1ms bins and produce movies of duration N-T = 250 ms (i.e. approximately corresponding to the average duration between two saccades). To avoid border effects, we will position a draw of the complete trajectory at random in the image space so that the sub-image is translated using the position given by the trajectory at each time step. The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space [Perrinet, 2015]. Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that a velocity of V=1 corresponds to a movement of one pixel per frame (i.e., per time bin).

To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames. If, on a specific pixel at that specific timestamp, the absolute value of this gradient exceeds a threshold, an event is generated. The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive. This signed threshold value is then subtracted from the residual gradient image. When applied to the whole movie, the event stream is then similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion-task}-(c)). The goal here is to infer the correct motion solely by observing these events.
